Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-28 paper/paper.tex
Clarification
Fixing intro

Broadly speaking, abstract abstract machines and abstract definitional interpreters are
different forms of abstract interpreters. They are obtained by applying a combination 
of abstractions to their concrete counterparts, abstract machines and definitional 
interpreters, respectively. An interesting question, and the subject of this paper, is 
how we can derive an abstract semantic artifact from the respective other one.

In the concrete world, the relations among reduction semantics, abstract machines,
definitional interpreters, and monadic interpreters have been intensively studied by
Danvy and his collaborators \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202,
danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, 
Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
The concrete abstract machines implement structural operational semantics in 
continuation-passing style, where the reduction contexts are defunctionalized continuations.
One can derive definitional interpreters by refunctionalizing the reduction contexts of 
abstract machines, and by defunctionalizing the higher-order functions, one may obtain 
abstract machines in the reverse direction.

In the domain of abstract semantic artifacts, by contrast, the relations between small-step
abstract abstract machines and big-step abstract definitional interpreters, as well as the 
question of deriving one from the other, are not well studied.
One of the fundamental differences between concrete semantic artifacts and abstract semantic
artifacts is a sound latter one must tell you facts about all possible execution paths, and 
thus nondeterminism is introduced. In addition to ensuring termination, abstract semantic
artifacts are usually equipped with a cache of reachable states, the cache is
updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
In this functional pearl, we make a first step to answering the question of interderiving
abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI 
by presenting a series of syntactic transformations on the program from small-step 
abstract abstract machines to big-step abstract definitional interpreters.

In addition, the abstract abstract machines with unbounded stack naturally correspond
to abstract definitional interpreters. We show that after refunctionalizing an AAM with 
unbounded stack, and with the proper caching algorithm, the pushdown control-flow analysis 
can be obtained.

>>>>

Broadly speaking, abstract abstract machines and abstract definitional interpreters are
different forms of abstract interpreters. They are obtained by applying a combination
of abstractions to their concrete counterparts, abstract machines and definitional
interpreters, respectively. An interesting question, and the subject of this paper, is
how one can derive an abstract semantic artifact given the counterpart.

In the concrete world, the relations among reduction semantics, abstract machines,
definitional interpreters, and monadic interpreters have been intensively studied by
Danvy and his collaborators \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202,
danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional,
Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
These concrete abstract machines implement structural operational semantics in 
continuation-passing style, where the reduction contexts are defunctionalized continuations.
One can derive definitional interpreters by refunctionalizing the reduction contexts of
abstract machines, and by defunctionalizing the higher-order functions, one may obtain
abstract machines in the reverse direction.

In the domain of abstract semantic artifacts, by contrast, the relations between small-step
abstract abstract machines and big-step abstract definitional interpreters, as well as the
question of deriving one from the other, are not well studied.
One of the fundamental differences between concrete semantic artifacts and abstract semantic
artifacts is that a sound instance of the latter must tell you facts about all possible execution paths, and
thus introduce nondeterminism. In addition to ensuring termination, abstract semantic
artifacts are usually equipped with a cache of reachable states; the cache is
updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
In this functional pearl, we make a first step in answering the question of interderiving
abstract semantics artifacts, and provide a constructive answer to relate AAMs and ADIs
by presenting a series of syntactic transformations on the program from small-step
abstract abstract machines to big-step abstract definitional interpreters.

In addition, those abstract abstract machines with unbounded stacks naturally correspond
to abstract definitional interpreters. We show that after refunctionalizing an AAM with an
unbounded stack (and utilizing the proper caching algorithm), the pushdown control-flow analysis
can be obtained.
